Elephant Seal Age JAE Intraspecific Variation

### 2023-12-07 paper/paperv3.qmd
Moved supplemental table to end

```{r}
#| tbl-cap: str_glue("The results from breeding probability analyis. The random effect variance of animal id is {VarCorr(breeding_mod, digits = 3)[1]} and year was VarCorr(breeding_mod, digits = 3)[2].")
#| eval: true

breed_mod_summary$coefficients %>% 
  as_tibble() %>% 
  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
         across(c(Estimate, `Std. Error`, `z value`), 
                \(x) format_signif(x, digits = 3)),
         P_value = format_pval(`Pr(>|z|)`, digits = 2)) %>%
  select(-`Pr(>|z|)`) %>% 
  relocate(Parameter) %>% 
  knitr::kable()
```

@@

breed_nonbreed <- sealdat %>% 
  group_by(age) %>% 
  summarize(n_observed = sum(!is.na(observed)), 
            perc_breed = sum(observed == "Breeder") / n_observed, 
            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
  filter(age < 22) %>%
  ggplot(aes(x = age, y = perc_breed)) + 
  geom_ribbon(aes(x, ymin = conf.low, ymax = conf.high),
              breed_pred,
              inherit.aes = FALSE,
              fill = "#829BD1",
              alpha = 0.25) +
  geom_line(aes(x, predicted),
            breed_pred,
            inherit.aes = FALSE,
            color = "#829BD1",
            linewidth = 1) +
  geom_point() + 
  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
  scale_y_continuous(labels = scales::percent, limits = c(.4, 1)) + 
  labs(x = "Maternal Age (Years)", y = "% Breeders (Mean ± SE)") + 
  theme_few() + 
  theme(axis.title = element_text(size = 13))

@@

```{r}
#| label: fig-breeding_bias
#| fig-width: 9

year_coef <- as_tibble(ranef(breeding_mod)$year, 
                       rownames = "year") %>% 
    mutate(year = as.numeric(year)) %>% 
    rename(coef = `(Intercept)`)

animal_coef <- as_tibble(ranef(breeding_mod)$animalID, 
                         rownames = "animalID") %>% 
    mutate(animalID = as.numeric(animalID)) %>% 
    rename(coef = `(Intercept)`)

year_bias <- sealdat %>% 
  left_join(year_coef, by = "year") %>% 
  mutate(age = factor(age)) %>% 
  ggplot(aes(age, coef)) +
  geom_boxplot() +
  geom_hline(yintercept = 0, color = "#829BD1", linetype = "dashed") +
  scale_x_discrete("Maternal Age (Years)", breaks = seq(5, 20, by = 5)) +
  ylab("Random effect of year") +
  theme_few()

animalID_bias <- sealdat %>% 
  left_join(animal_coef, by = "animalID") %>% 
  mutate(age = factor(age)) %>% 
  ggplot(aes(age, coef)) +
  geom_boxplot() +
  geom_hline(yintercept = 0, color = "#829BD1", linetype = "dashed") +
  scale_x_discrete("Maternal Age (Years)", breaks = seq(5, 20, by = 5)) +  
  ylab("Random effect of individual") +
  theme_few()

plot_grid(breed_nonbreed, year_bias, animalID_bias, nrow = 1)

@@

```{r}
#| label: fig-repro_age
#| fig-cap: "Pup survival decreases slightly, but survival and recruitment are not significantly affected by maternal age."
#| fig-width: 4.5

@@

## NOTE: REMOVED RANDOM EFFECT OF INDIVIDUAL, OTHERWISE SINGULAR FIT (HYPOTHESIS: 
## INSUFFICIENT SAMPLE SIZES PER INDIVIDUAL)

@@

surv_recruit_summary %>% 
  ggplot(aes(x = age, y = mean, color = metric, fill = metric)) + 
  geom_ribbon(aes(ymin = mean_lwr, ymax = mean_upr, alpha = metric),
              surv_recruit_pred,
              color = NA) +
  geom_pointrange(aes(ymin = low, ymax = high)) +
  geom_line(aes(linetype = metric),
            surv_recruit_pred) +
  scale_y_continuous(labels = scales::percent) + 
  scale_color_manual(values = c("#829BD1", "#E9B668")) + 
  scale_fill_manual(values = c("#829BD1", "#E9B668")) + 
  scale_linetype_manual(values = c("solid", "dashed")) +
  scale_alpha_manual(values = c(0.25, 0.1)) +
  expand_limits(y = c(0, 0.5)) +
  labs(x = "Maternal Age (Years)", y = "Pup Survival or Recruitment") +
  theme_few() + 
  theme(legend.position = c(0.1,0.97), 
        legend.justification = c(0,1), 
        legend.title = element_blank(), 
        legend.direction = "horizontal", 
        axis.title = element_text(size = 13))

@@

>>>>

@@

breed_nonbreed <- sealdat %>% 
  group_by(age) %>% 
  summarize(n_observed = sum(!is.na(observed)), 
            perc_breed = sum(observed == "Breeder") / n_observed, 
            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
  filter(age < 22) %>%
  ggplot(aes(x = age, y = perc_breed)) + 
  geom_ribbon(aes(x, ymin = conf.low, ymax = conf.high),
              breed_pred,
              inherit.aes = FALSE,
              fill = "#829BD1",
              alpha = 0.25) +
  geom_line(aes(x, predicted),
            breed_pred,
            inherit.aes = FALSE,
            color = "#829BD1",
            linewidth = 1) +
  geom_point() + 
  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) + 
  labs(x = "Maternal Age (Years)", y = "% Breeders (Mean ± SE)") + 
  theme_few() + 
  theme(axis.title = element_text(size = 13))

@@

@@

```{r}
#| label: fig-repro_age
#| fig-cap: "Pup survival and recruitment are not significantly affected by maternal age. Points are mean values for maternal age. Error bars show SE."
#| fig-width: 4.5

@@

@@

surv_recruit_summary %>% 
  ggplot(aes(x = age, y = mean, color = metric, fill = metric)) + 
  geom_ribbon(aes(ymin = mean_lwr, ymax = mean_upr, alpha = metric),
              surv_recruit_pred,
              color = NA) +
  geom_pointrange(aes(ymin = low, ymax = high)) +
  geom_line(aes(linetype = metric),
            surv_recruit_pred) +
  scale_y_continuous(labels = scales::percent) + 
  scale_color_manual(values = c("#829BD1", "#E9B668")) + 
  scale_fill_manual(values = c("#829BD1", "#E9B668")) + 
  scale_linetype_manual(values = c("dashed", "dashed")) +
  scale_alpha_manual(values = c(0.25, 0.1)) +
  expand_limits(y = c(0, 0.5)) +
  labs(x = "Maternal Age (Years)", y = "Pup Survival or Recruitment") +
  theme_few() + 
  theme(legend.position = c(0.1,0.97), 
        legend.justification = c(0,1), 
        legend.title = element_blank(), 
        legend.direction = "horizontal", 
        axis.title = element_text(size = 13))

@@

### Supplemental Material

Supplemental Table 1: Statistics for breeding success model.

```{r}
#| label: tbl-breed-cap

tbl_breed_cap <- str_glue("The results from breeding probability analyis. The random effect variance of animal id is {format_signif(VarCorr(breeding_mod)$animalID, 3)} and of year was {format_signif(VarCorr(breeding_mod)$year, 3)}.")

```{r}
#| label: tbl-breed-success
#| tbl-cap: !expr "tbl_breed_cap"

breed_mod_summary$coefficients %>% 
  as_tibble() %>% 
  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
         across(c(Estimate, `Std. Error`, `z value`), 
                \(x) format_signif(x, digits = 3)),
         `p value` = format_pval(`Pr(>|z|)`, digits = 2)) %>%
  select(-`Pr(>|z|)`) %>% 
  relocate(Parameter) %>% 
  knitr::kable()
```
